\LaTeX\ Author Guidelines for 3DIMPVT Proceedings

### 2012-06-19 3dimpvt/paper.tex
just waiting for some images

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with higher localization
errors and greater variance in camera locations. The paper is
organized as follows. Section 2 provides an overview of the backpack
modeling system and data processing pipeline from which input data and
examples used throughout this paper originate. Section 3 describes the
general problem of texture mapping and examines two simple mapping
approaches and their problems.  Section 4 demonstrates previous
attempts at localization refinement, and presents our new approach to
image alignment. Section 5 combines our image alignment algorithm with
a simple image selection approach, as well as with a new method for image selection that
takes advantage of optimal projections. Section 6 contains results and
conclusions.

@@

\subsection{Data Acquisition Hardware}
The backpack modeling system that captured our data contains five 2D
laser range scanners, two cameras, an orientation sensor, and an
IMU. The laser scanners are mounted orthogonally and have a 30-meter
range and a 270$^{\circ}$ field of view. The two cameras are equipped
with fisheye lenses, resulting in a 180$^{\circ}$ field of view, and
are mounted with one facing left and one facing right. These cameras take images at the rate of 5 Hz. The orientation
sensor provides orientation parameters at a rate of 180 Hz. The IMU
provides highly accurate measurements of all 6 DOF at 200 Hz, and is
used as a ground truth reference.

With the backpack actively scanning, the human operator wearing it
then takes great care to walk a path such that every wall in the
desired indoor environment is passed and scanned lengthwise at least
once.

\subsection{Environment Reconstruction}
Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is first localized over its data
collection period, and a 3D point cloud of the surrounding environment
is constructed based on the laser scanner readings relative to the backpack. Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This model,
consisting of multiple 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's camera, can be considered the
input to our texture mapping problem.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square
tiles, generally about 5 pixels across, and picking an image to texture that tile with. In order to
do this, we must first gather a list of candidate images for each tile
$t$. These candidate images must contain all four corners of $t$,
which we can quickly check using the projection method
above. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to the center of $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, our section, and other planes, all in
world coordinates.

@@

A scoring function to maximize these two criteria is as follows:

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.png}
  \caption{For each tile, we want to select an image such that the
    function $abcdefg/awerw$ is maximized}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{directMapping.png}
  \caption{The result of direct texture mapping based on locally optimized textures.}
  \label{fig:directMapping}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{caching.png}
  \caption{The result of adding a caching system to locally optimized textures.}
  \label{fig:caching}
\end{figure}

@@

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at a head-on angle and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting as well
as blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections.

@@

When images are taken of a plane from arbitrary overlapping positions,
they are related by homography \cite{hz}. Thus, existing
homography-based image mosaicing algorithms are applicable
\cite{brown2007automatic}. However, errors can compound when long
chains of images are mosaiced together using these approaches. For
example, a pixel in the $n$th image in the chain must be translated
into the first image's coordinates by multiplying by the $3\times3$
matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these homography
matrices is propagated to all further images until the chain is
broken. For some chains of images this can happen almost immediately
due to erroneous correspondence matches and the resulting image mosaic
is grossly misshapen.

Figure \ref{fig:mosaic} shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane. This image was generated after
careful hand tuning. Many planes that had fewer features simply
failed. This led to the conclusion that image mosaicing is not a
robust enough solution for reliably texture mapping our dataset.

@@

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom. Previous work has attempted to refine camera
matrices by solving a non-linear optimization problem
\cite{liu2010indoor}. This process is specific to the backpack system
which generated our dataset, as it must be run during backpack
localization\cite{liu2010indoor,chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem shown in
Figure \ref{fig:graph}. In our new approach, we also refine the
placement of images using image correspondences. However, we do so in
two dimensions on the plane whereas this previous work did so over all
6 degrees of freedom. Refining in two dimensions on the plane is less
flexible in that it does not address projection errors, however, it
provides noticeable benefits while avoiding the error propagation
problem.

\section{Our Approach}
Our approach begins with the projection of all our images onto
separate copies of our plane, such that no projected data is covered
and lost. Tis is done by discretizing the plane into an arbitrary
density of points, and projecting each point into the image plane
using the camera matrix $P_i$. We then perform a series of
preprocessing steps to rotate and shift each image's projection such
that its overlapping regions with other projections are as seamless as
possible. From here, we assign textures to our planes using one of two
methods, dependent on whether they are wall planes or ceiling/floor
planes. The reasoning for this is explained in section
\ref{sec:whydifferent}. Lastly, we apply two post-processing steps to
smooth out any remaining seams.

\subsection{Hough Transforms}
Our first step is to use Hough transforms to rotate the projected
images such that any strong near-vertical features are made completely
vertical. This is effective for indoor modeling, since many indoor
scenes contain strong vertical lines from features such as doors, wall
panels, or rectangular frames.

\subsection{Localization Refinement}
Our next step is to try and fix misalignment between overlapping
images.  We do this by fixing the locations in 2D. First we find
corresponding points between all pairs of overlapping images using
SIFT matches \cite{lowe1999object}. An illustration of this is given
in Figure \ref{fig:matches}. The SIFT matches allow us to determine
feature-based $x$ and $y$ distances between two images on the plane,
which we can compare against the $x$ and $y$ distances provided by the
camera matrices. With these data points, we can reevaluate where each
image should be projected.

\subsubsection{Robust SIFT Distances using RANSAC}

First we need to ensure that the SIFT-based distances are
reliable. Since SIFT matches frequently include outliers, the RANSAC
framework \cite{fischler1981random} is used for a robust estimate of
the $x$ and $y$ distances between the two images. The RANSAC framework
will attempt to build a consensus among the SIFT matches about what
the true $x$ and $y$ distance is between the two images while ignoring
the influence of outliers that may skew the results. The framework
handles the consensus-building machinery, and only requires that two
functions be specified: the fitting function and the distance
function. These functions are called for random subsets of the SIFT
matches until the best set of inliers is found. For this application,
the fitting function simply finds the average distance between
matches. If the matches are exactly correct and the image is frontal
and planar then the distances for various SIFT feature matches should
be the same. Our distance function for a pair of points is the
difference between those points' SIFT match distance and the average
distance computed by the fitting function. We specified a 10 pixel
outlier threshold to the framework. This means that a SIFT match is
labeled as an outlier if its $x$ and $y$ distance is not within 10
pixels of the average distance computed by the fitting function.

@@

There are a total of $M^{2}$ possible pairs of images, however we can
only measure distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we solve a least squares problem ($\textrm{min}_{\beta}
||\beta X - y||_2^2 $) to estimate the correct location of the images
on the plane. The vector $\beta$ of unknowns are the correct $x$ and
$y$ locations of each image on the plane from $1 \dots M$. Finding the
correct $x$ and $y$ locations are independent from one another so we
will only consider the $x$ locations:

@@

Finally, the observation vector $y$ is constructed using the
SIFT-based distances generated earlier in the matching stage. The
distances are denoted as $d_1 \dots d_N$ for $N$ SIFT-based
distances. The last element in the observation vector is the original
location of the first image determined by the localization algorithm.

@@

The $\beta$ that minimizes $||\beta X - y||_2^2$ tells us a set of
image locations on the plane that best honors all the SIFT-based
distance measurements between images. A similar problem is solved for
the $y$ dimension. In practice there are often cases where there is a
break in the chain of images, meaning that no SIFT matches were found
between one segment of the plane and another. In this case we add rows
to the $X$ matrix and observations to the $y$ vector that contain the
original $x$ and $y$ distance estimates generated by the localization
algorithm. Another way to do this is to add rows for all neighboring
pairs of images and solve a weighted least squares problem where the
SIFT distances are given a higher weight i.e. 1, and the distances
generated by the localization algorithm are given a smaller weight
i.e. 0.01.

At this point we now have our target plane associated with a list of
images as well as their rotated and shifted projections upon the
plane. We are ready to begin assigning these projections to areas of
our plane for texturing.

\subsection{Ceiling/Floor Texturing vs. Wall Texturing}
\label{sec:whydifferent}
The backpack modeling system, as described in Section
\ref{sec:backpackHardware}, was designed to prioritize for texturing
walls. As mentioned previously, the backpack contains two fisheye
cameras, one aimed directly to the left of the backpack operator, and
one aimed directly to the right. Thus, in order to texture the ceiling
above where the operator walked, images from both cameras must be
combined, with their boundaries meeting overhead. Unlike walls, where
one image generally spans the entire height of a wall, two images are
required to span the width of a ceiling. Even more problematic, the
floor beneath the backpack is physically obscured by the operator's
body, and thus must be textured by images taken from behind the
operator or from the side. This problem of distant cameras at poor
angles is compounded by the fact that the operator's data acquisition
path attempts to fully scan each wall, but not necessarily the
entirety of a floor or ceiling. This means that large swaths of floors
and ceilings are never walked directly over or under, and thus must be
textured by sidelong images taken from further distances and at
oblique angles. These effects are shown in the following figure.

In short, these factors mean that wall planes are generally associated
with a multitude of upright, uniform images, from a relatively close
distance and taken head-on at regular intervals, while floors and
ceilings are associated with images taken from all manner of
orientations and distances. This major difference suggests we employ
two strategies, one tailored for each scenario.

\subsection{Texturing Walls}
Because wall images tend to all be fairly good candidates for
texturing due to their orientation and location, our goal is less
about determining the best images, and more about selecting a set of
images that together form the most seamless texture.

\subsubsection{Cost Functions}

In order to objectively decide which set of images accomplishes this,
we need to decide on an overall cost function. As our goal is to
reduce the seams that occur where two images meet, a simple cost
function is the sum of squared pixel differences in overlapping
regions between every pair of images. This encourages image boundaries
to occur in featureless areas, such as bare walls, on in areas where
images are very well aligned.

Another possible cost function is edge energy, i.e. the sum of the
smoothed gradient of the blended seam regions. Minimizing this
encourages image boundaries to be placed in featureless areas even
more than the first cost function. For the results shown throughout
this paper, the first cost function was used, as the regular
occurrence of featureless areas was not guaranteed in our datasets.

@@

Now that we have a cost function defined, we need to mechanically
select the set of images for which the overall cost function is
minimized. Since we aim to cover the entirety of our plane, our
problem is to minimally cover a polygon(our plane) using other
polygons of arbitrary geometry (our image projections), with the added
constraint of minimizing our cost function. This is a complex problem,
though it can be generalized to a number of graph coverage problems,
as depicted in figure X. Returning again to the optimality of our
situation when texturing walls however, we can make a quick
simplification for the sake of efficiency. Given that our wall-texture
candidate images were all taken from a head-on angle, and assuming
only minor rotations were made earlier, we can reason that their
projections onto the plane are approximately rectangular. By
discarding any excess texture and cropping them all to be rectangular,
our problem becomes the conceptually simpler problem of filling a
polygon with rectangles, such that the sum of all edge costs between
each pair of rectangles is minimal. An additional benefit of working
with rectangular units is that borders between any two distinct images
in our final texture will be either horizontal or vertical. Since most
environmental features inside buildings are horizontal or vertical,
any seams in our texture will intersect them miniimally and be less
noticeable.

By taking one more shortcut we can simplify our problem down even
further. Since care is taken such that our wall images contain the
floor-to-ceiling range of walls, we generally do not have any wall
images such that one should be projected vertically above the
other. In essence, we need only to ensure horizontal coverage of our
planes, as our images provide full vertical coverage. We can thus
construct a Directed Acyclic Graph (DAG) from the images, again with
edge costs defined by our cost function above, and solve a simple
shortest path problem to find an optimal subset of images with regard
to the cost functions.

\begin{figure}
  \centering
  \includegraphics[width=3in]{DynProg.pdf}
  \caption{Image Selection is done by constructing a graph of sorted
    images. Then we solve a shortest path problem where the edge
    weights represent the cost of a seam between two overlapping
    images.}
  \label{fig:DynProg}
\end{figure}

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by a chosen image, we are left with a hole where no
image was chosen to texture. Rather than reverting to a 2D-coverage
problem, we can elect to simply fill the hole by selecting images in a
greedy fashion, with respect to the same cost function.

At this point, we have mapped every location on our plane to an image
or images that were selected. Before applying these textures, we will
examine how to accomplish the same mapping for ceilings and floors.

\subsection{Texturing Ceilings and Floors}

Because the images associated with ceilings and floors are often
captured by cameras at sidelong angles, we must be conservative when
we texture map them. Since localization errors compound over distance,
we often have candidate images that project well on areas near the
camera itself, but much more poorly at the greater distances caused by
higher angles. Using entire images to minimize seams is beneficial for
wall texturing, but it becomes detrimental on ceilings and floors,
since the quality of usable texture varies greatly across any one
image. As evident in the following comparison, ceiling and floor
images often have a very low amount of optimal texture relative to the
projection size. This factor suggests that we should segment our
images in some way, as many of our images contain both desired as well
as undesired portions.

Such segmentation implies smaller units of texture, which often leads
to more images being chosen and unfortunately, more boundaries between
different images as a result. This goes strongly against our approach
for texturing walls, in which we attempted to minimize the number of
images chosen. Nevertheless, boundaries on floors and ceilings are
less problematic, since floors and ceilings tend to have fewer
noticeable features.

As shown in the above figure, we define a tile length $l1$ as well as
an overlap length $l2$, usually about 5 pixels for each. We then
discretize our plane into rectangular tiles $l1$x$l1$ in size. Each
tile is then associated with a square texture of size
$(l1+2l2)$x$(l1+2l2)$, such that it shares a large overlapping region
with its neighboring tiles, allowing for blending afterwards. As in
wall texturing, rectangles, rather than triangles, are preferred for
the high prevalence of horizontal and vertical features, generally
light fixtures and floor and ceiling tiles.

For the actual process of image association, we can reuse the caching
method shown earlier. This method will balance selection of the best
texture for each tile with some amount of image reuse. Coupled with
image rotation and shifting, as well as the usage of rectangles
instead of triangles, we end with a much more seamless texture, even
without any blending applied yet, as seen below.

\subsection{Blending}
Though our preprocessing steps and image selection attempts above
attempted to minimize all mismatches between images, there are often
cases, where due to geometry or projection inaccuracies, we simply
have unavoidable discontinuities. These can however be partially
treated by applying alpha blending over our seams.  Whether the units
we are blending are rectangularly-cropped images for walls, or
rectangular units for ceilings/floors, we can apply the same blending
procedure.

Our alpha blending technique is widely used, and blends pixels
linearly across overlapping regions. For each pixel within an
overlapping region, we simply multiply its intensity (by a factor from
0 to 1), based on its distance from the image's border, with a cap set
at some maximum blending distance. In this way, we interpolate between
two overlapping images, providing a gradual transition between
them. Increasing the overlapping and blending distance results in a
smoother transition, but can also result in more blurriness. The final
output is shown in Figure X.

>>>>

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with higher localization
errors and greater variance in camera locations. The paper is
organized as follows. Section 2 provides an overview of the backpack
modeling system and data processing pipeline from which input data and
examples used throughout this paper originate. Section 3 describes the
general problem of texture mapping and examines two simple mapping
approaches and their problems.  Section 4 demonstrates previous
attempts at localization refinement, and presents our new approach to
image alignment. Section 5 combines our image alignment algorithm with
a simple image selection approach, as well as with a new method for
image selection that takes advantage of optimal images. Section 6
contains results.

@@

\subsection{Data Acquisition Hardware}
The backpack modeling system that captured our data contains five 2D
laser range scanners, two cameras, an orientation sensor, and an
IMU. The laser scanners are mounted orthogonally and have a 30-meter
range and a 270$^{\circ}$ field of view. The two cameras are equipped
with fisheye lenses, reaching an approximately 180$^{\circ}$ field of
view, and are mounted with one facing left and one facing right. These
cameras take images at the rate of 5 Hz. The orientation sensor
provides orientation parameters at a rate of 180 Hz. The IMU provides
highly accurate measurements of all 6 DOF at 200 Hz, and is used as a
ground truth reference.

With the backpack actively scanning, the human operator wearing it
takes great care to walk a path such that every wall in the desired
indoor environment is passed and scanned lengthwise at least once.

\subsection{Environment Reconstruction}
Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is first localized over its
data collection period, and a 3D point cloud of the surrounding
environment is constructed based on the laser scanner readings
relative to the backpack \cite{chen2010indoor}. Approximate normal vectors for
each point in the point cloud are then calculated by gathering
neighboring points within a small radius and processing them through
principal component analysis. These normal vectors allow for the
classification and grouping of adjacent points into structures such as
walls, ceilings, floors, and staircases. A RANSAC algorithm is then
employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model \cite{sanchez2012point}. This model,
consisting of multiple 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's camera, can be considered the
input to our texture mapping problem.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and picking an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture will be
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
will intersect them miniimally and be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check using the projection method
above. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to the center of $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, our section, and other planes, all in
world coordinates.

@@

A scoring function to maximize these two criteria is shown in figure \ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{For each tile, we find the image that minimizes $d$ and $\alpha$ by maximizing the following equation: $(1/d \cdot (-1 \cdot C) \cdot N)$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_naive.jpg}
  \caption{The result of direct texture mapping based on locally
    optimized textures.}
  \label{fig:directMapping}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_cache_full.jpg}
  \caption{The result of adding a caching system to locally optimized textures.}
  \label{fig:caching}
\end{figure}

@@

Rather than perform camera or image adjustments across many thousands
of images, we choose to work with the more limited set of images
corresponding to those chosen by the direct mapping approach, without
caching. This set of images constitutes a good candidate set for
generating a final seamless texture, since it meets three important
criteria. First, this set of images contains at least one image that
covers each tile on our plane. Thus, unless extreme changes in
localization occur, we can ensure that there will be no holes in our
final texture. Second, since images are all selected according to the
same scoring function in figure \ref{fig:scoringFunction}, we know
that our images are all taken at a head-on angle and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting as well
as blending between them. With this set of images, we now examine two
approaches towards refining and combining their projections, before demonstrating our own.

@@

When images of a plane are taken from arbitrary overlapping positions,
they are related by homography \cite{hz}. Thus, existing
homography-based image mosaicing algorithms are applicable
\cite{brown2007automatic}. However, errors can compound when long
chains of images are mosaiced together using these approaches. For
example, a pixel in the $n$th image in the chain must be translated
into the first image's coordinates by multiplying by the $3\times3$
matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these homography
matrices is propagated to all further images until the chain is
broken. For some chains of images this can happen almost immediately
due to erroneous correspondence matches and the resulting image mosaic
is grossly misshapen.

Figure \ref{fig:mosaic} shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane. This image was generated after
careful hand tuning as well. Many planes that had fewer features simply
failed outright. This leads to the conclusion that image mosaicing is not a
robust enough solution for reliably texture mapping our dataset.

@@

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem attempted to refine camera
matrices by solving a non-linear optimization problem
\cite{liu2010indoor}. This process is specific to the backpack system
which generated our dataset, as it must be run during backpack
localization\cite{liu2010indoor,chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem shown in
Figure \ref{fig:graph}. In our new approach, we also refine the
placement of images using image correspondences. However, we do so in
two dimensions on the plane whereas this previous work did so over all
6 degrees of freedom. Refining in two dimensions on the plane is less
flexible in that it does not address projection errors, however, it
provides significant benefits while avoiding the error propagation
problem.

\subsection{Improved Localization through 2D Matching}
Our approach begins with the projection of all our images onto
separate copies of our plane, such that no projected data is covered
up and lost. This is done in the same way as the approaches in section
3. We then perform Hough transforms to rotate these projections
similarly, and then shift each image's projection such that its
overlapping regions with other projections are as seamless as
possible.

\subsection{Hough Transforms}
Our first step is to use Hough transforms to adjust the rotation of
projected images. Rather than try and match the orientation of
features in each image, we simply apply rotations such that any strong
near-vertical features are made completely vertical. This is effective
for indoor modeling, since strong features in indoor scenes usually
contain vertical lines corresponding to doors, wall panels,
rectangular frames, etc. If features in the environment are not
vertical, or there are few features overall, this step should be
skipped.

\subsection{SIFT Feature Matching}
Our next step is to try and fix misalignment between overlapping
images. We do this by first searching for corresponding points between
all pairs of overlapping images using SIFT feature matching
\cite{lowe1999object}. An illustration of this is given in Figure
\ref{fig:matches}. The SIFT matches allow us to determine
feature-based $x$ and $y$ distances between two images on the plane,
which will allow us to reevaluate where each image should be projected.

\subsubsection{Robust SIFT Distances using RANSAC}
Since indoor environments often contain repetitive features, such as floor tiles or doors, we need to ensure that our SIFT-based distances are
reliable. In order to diminish the effect of incorrect matches and outliers, the RANSAC
framework \cite{fischler1981random} is used for a robust estimate of
the $x$ and $y$ distances between the two images. The RANSAC framework
attempts to build a consensus among the SIFT matches between every pair of images, and tries to generate the horizontal and vertical distance between them while ignoring
the influence of outliers that would skew the results. The framework
handles the consensus-building machinery, and only requires that two
functions be specified: the fitting function and the distance
function. These functions are called for random subsets of the SIFT
matches until the best set of inliers is found. For this application,
the fitting function simply finds the average distance between
matches. If the matches are exactly correct and the image is frontal
and planar then the distances for various SIFT feature matches should
be the same. Our distance function for a pair of points is the
difference between those points' SIFT match distance and the average
distance computed by the fitting function. We specified a 10 pixel
outlier threshold to the framework. This means that a SIFT match is
labeled as an outlier if its horizontal or vertical distance is not within 10
pixels of the average distance computed by the fitting function.

@@

There are a total of $M^{2}$ possible pairs of images, though we only generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem ($\textrm{min}_{\beta}
||\beta X - y||_2^2 $) to estimate the correct location of the images
on the plane. The vector $\beta$ of unknowns represent the correct $x$ and
$y$ locations of each image on the plane from $1 \dots M$. The optimal $x$ and $y$ locations are calculated in the same way, so we
will only consider the $x$ locations here:

@@

Finally, the observation vector $y$ is constructed using the
SIFT-based distances generated earlier in the matching stage. The
distances are denoted as $d_1 \dots d_N$ for $N$ SIFT-based
distances. The last element in the observation vector is the location of the first image determined by its original localization.

@@

The $\beta$ that minimizes $||\beta X - y||_2^2$ results in a set of
image locations on the plane that best honors all the SIFT-based
distance measurements between images. In practice there are often cases where there is a
break in the chain of images, meaning that no SIFT matches were found
between one segment of the plane and another. In this case we add rows
to the $X$ matrix and observations to the $y$ vector that contain the
original $x$ and $y$ distance estimates generated by the localization
algorithm. Another way to do this is to add rows for all neighboring
pairs of images and solve a weighted least squares problem where the
SIFT distances are given a higher weight i.e. 1, and the distances
generated by the localization algorithm are given a smaller weight
i.e. 0.01.

After completing this same process for the $y$ dimension as well, and making the resultant shifts, our image projections should overlap and match eachother with far greater accuracy.

\section{Improved Texture Mapping with Localization and Blending}
Now that our images have been relocalized for much greater accuracy relative to eachother, we can revisit the cached mapping approach from section 3.2. Since we now have higher confidence in the quality of image boundaries, we can also explore an alternate method that greatly reduces seams in cases where we have an abundance of optimal images.

\subsection{Localized Mapping with Caching}
In figure \ref{fig:shifted}, we see the same wall from section 3,
texture mapped using the caching method, both without and with
localization for comparison.

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_cache_full_shifted.jpg}
  \caption{Localization refinement results in signifcantly fewer
    discontinuities in the final texture.}
  \label{fig:shifted}
\end{figure}

It should be evident that our localization adjustments resulted in a
major improvement. A great strength of our tiling approach for image
selection is that it runs regardless of plane geometry or orientation,
and can handle images of all types. For example, here is the approach
as run on a ceiling and floor plane, both of which contain a multitude
of non-optimal images due to the camera locations on the backpack.

% \begin{figure}
%   \centering
%   \includegraphics[width=3in]{localized_caching_ceiling.png}
%   \caption{Using the graph-based localization refinement algorithm
%   from [11] suffers from the problem of compounding errors. }
%   \label{fig:graph}
% \end{figure}

Given that the backpack system's hardware and the data collection
process do strive to collect optimal images for walls however, we can
be less reserved when texturing walls, and simply fall back to this
more robust approach as needed.

\subsection{Texture Mapping with Seam Minimization}
The cached tile approach selected images based on individual quality
and neighboring quality, in an effort to reduce seams between
tiles. In a more optimal case, as the backpack system provides for
walls, images all have near-rectangular projections and thus a great
score according to the scoring function in figure
\ref{fig:scoringFunction}. Thus, we can reason that rather than
selecting the set of best images, since all images are good, we should
instead select the best set of images, such that the selection results
in the cleanest texture possible. We will accomplish this by using
entire images where possible, defining a cost function between images,
and seeking to maximize our total cost.

\subsection{Occlusion Masking}
Before we proceed with using our images, we need to ensure that they
contain only content that should be mapped onto the target plane in
question. The tiling approach used previously only checked occlusion
for each tile as it was being textured, so we need to perform
occlusion checks over the entirety of each image, so we know which
areas are available for texture mapping. By performing the same
intersection tests used for tiles, we can determine whether individual
sections of an image are occluded.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, our image projections will remain largely rectangular. We can
thus recursively split each image into rectangular pieces, and perform
occlusion checks, similar to the process done for tiles earlier. To
actually occlude out rectangles, we simply remove their texture, as we
will ensure that untextured areas are never chosen for texture
mapping.

\subsubsection{Inter-Image Cost Functions}
In order to objectively decide which set of images results in the
cleanest texture, we need a cost function such that we can evaluate
the visibiilty of seams between images in our set. A straightforward
cost function that accomplishes this is the sum of squared pixel
differences in overlapping regions between all pairs of
images. Minimizing this cost function encourages image boundaries to
occur in featureless areas, such as bare walls, as well as in areas
where images match extremely accurately.

Another possible cost function is overall edge energy, i.e. the sum of
the smoothed gradient over seams. Minimizing this encourages image
boundaries to be placed in featureless areas even more than the first
cost function. For the results shown throughout this paper, the first
cost function was used, as the regular occurrence of featureless areas
was not guaranteed in our datasets, and our image matches do in fact
match at the pixel level in many cases.

@@

Now that we have a cost function defined, we mechanically select the
set of images for which the overall cost function is minimized. Since
we aim to cover the entirety of our plane, our problem is to minimally
cover a polygon(our plane), using other polygons of arbitrary geometry
(our image projections), with the added constraint of minimizing our
cost function between chosen images. This is a complex problem, though
we can take a number of steps to simplify it. Returning again to the
optimality of our situation when texturing walls however, we can make
a quick simplification for the sake of efficiency. Given that our
wall-texture candidate images were all taken from a head-on angle, and
assuming only minor rotations were made during localization
refinement, we can reason that their projections onto the plane are
approximately rectangular. By discarding the minor excess texture and
cropping them all to be rectangular, our problem becomes the
conceptually simpler problem of filling a polygon with rectangles,
such that the sum of all edge costs between each pair of rectangles is
minimal. We thus also retain axis-aligned image boundaries, along with
their advantages explained in section 3.

By taking one more shortcut we can simplify our problem down even
further. Since this approach is already assuming optimal images, we
know it will be applied only on walls, which are the focus of the
backpack modeling system. For wall planes, care is taken so that
images contain the entirety of their floor-to-ceiling range, and thus
we do not have wall images such that one should be projected
vertically above the other. In essence, we need only to ensure
horizontal coverage of our planes, as our images provide full vertical
coverage themselves. We can thus construct a Directed Acyclic Graph
(DAG) from the images, again with edge costs defined by our cost
function above, and solve a simple shortest path problem to find an
optimal subset of images with regard to the cost functions.

%\begin{figure}
%  \centering
%  \includegraphics[width=3in]{DynProg.pdf}
%  \caption{Image Selection is done by constructing a graph of sorted
%    images. Then we solve a shortest path problem where the edge
%    weights represent the cost of a seam between two overlapping
%    images.}
%  \label{fig:DynProg}
%\end{figure}

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by a chosen image, we are left with a hole where no
image was chosen to texture. Rather than reverting to a 2D-coverage
problem, we can elect to simply fill the hole by selecting images in a
greedy fashion with respect to the same cost function.

With this completed, we have now mapped every location on our plane to
at least one image, and have minimized the amount of images, as well
as the discontinuity between their borders. In the next section, we
will apply blending between images where they overlap, but for the
sake of comparison with the unblended tile caching method in section
5.1, we arbitrarily pick one image for texturing where images
overlap. The following images compare the tile caching method against
this seam minimization method.

%\begin{figure}
%  \centering
%  \includegraphics[height=2in]{compare_wall.png}
%  \caption{Planes are specified in 3D space by four corners $C_1$ to
%    $C_4$. Images are related to each plane through the camera
%    matrices $P_{1..M}$. }
  % \caption{The plane is specified in 3D space by the four corners
  % $C_1$ to $C_4$. Images are related to the plane through the camera
  % matrices $P_{1..M}$. For example, the 3D point X in the world
  % coordinate system is related to the image point x shown in the
  % figure as $x = P_i X$.}
%  \label{fig:compare_wall}
%\end{figure}

Though both methods provide quite accurate texturing thanks to the
localization refinement process, the seam minimization approach
results in fewer visible discontinuities, since it directly reduces
the cost of each image boundary, while the tile caching method uses a
scoring function that only approximates this effect. Furthermore, seam
minimization guarantees the best selection of images, while the tile
caching method may select images early on that turn out to be poor
choices once subsequent tiles have been processed. This can be a
problem when localization refinement does not work perfectly. Such an
effect can be seen in the following image.

In the context of the backpack modeling system, we apply the seam
minimization approach on walls, due to its superior output when
provided with optimal images. Floors and ceilings however, given their
suboptimal images which require segmentation, are textured using the
tile caching method.

\subsection{Blending}
Until now, blending of images has not been used, for the sake of clear
comparisons between texturing methods. We will now apply the same
blending process on our two final texturing methods: localization
refinement followed by either tile caching or seam minimization.

Although our preprocessing steps and image selections in either method
attempt to minimize all mismatches between images, there are cases,
where due to different lighting conditions or inaccuracies in geometry
or projection, where we simply have unavoidable discontinuities. These
can however be treated and smoothed over by applying alpha blending
over image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units.

For the tile caching method, we can ensure overlap by texturing a
larger tile than we really have. For example, for a tile $l_1 x l_1$
in size, we can make sure to associate it with a texture $(l_1 + l_2)
x (l_1 + l_2)$ in size. For the seam minimization method, we have
already ensured overlap between images. To enforce consistent blending
however, it is beneficial to add a required overlap distance while
solving the shortest path problem in section 5.2.3. If images do not
overlap in a region at least the length of this overlap distance, we
do not consider them to overlap at all. If images overlap in a region
greater than the overlap distance, we will only apply blending over an
area equal to the overlap distance.

Our alpha blending technique is widely used, and blends pixels
linearly across overlapping regions. For each pixel within such a
region, we simply scale its intensity (by a factor from 0 to 1),
based on its distance from the image's border, with a cap set at a fixed distance from the border. In this way, we interpolate between two
overlapping images, providing a gradual transition between
them. Increasing the blending distance results in a
smoother transition, but can also result in more blurriness. A blended and unblended version of each method is shown below for comparison.
